---
title: "ROSM: Open Street Map tiles in R"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/paleolimbot/rosm/branch/master/graph/badge.svg)](https://app.codecov.io/gh/paleolimbot/rosm?branch=master)
<!-- badges: end -->

Download and plot [Open Street Map](https://www.openstreetmap.org/), [Bing Maps](https://www.bing.com/maps), and other tiled map sources. Use to create high-resolution basemaps and add hillshade to vector based maps. Note that rosm uses base plotting and not ggplot2: for mapping in ggplot2, use [ggspatial](https://github.com/paleolimbot/ggspatial)::annotation_map_tile().

## Installation

The **rosm** package is [available on CRAN](https://cran.r-project.org/package=rosm), and can be installed using `install.packages("rosm")`.

## Example

```{r example, message=FALSE, warning=FALSE, results='hide', fig.show='hold'}
library(prettymapr)
library(rosm)

# specify a bounding box
altalake <- makebbox(50.1232, -122.9574, 50.1035, -123.0042)
prettymap(
  {
    # plot tiles (see also osm.raster() and bmaps.plot())
    osm.plot(altalake)
    osm.points(c(-122.9841, -122.9812), c(50.11055, 50.11765),
      pch = 15, cex = 0.6, col = "white"
    )
    osm.text(c(-122.9841, -122.9812), c(50.11055, 50.11765),
      labels = c("GC6", "GC2"), adj = c(-0.2, 0.5), cex = 0.7, col = "white"
    )
  },
  scale.label.col = "white"
)
```


## The Future?

Some other great projects have popped up in the last few years, such that **rosm** will probably be retired in the next year. Packages that depend on will continue to work, but it is likely that a new API will pop up under a different name.
